---
title: "p8105_hw6_ym2813"
author: "Anna Ma"
date: "12/2/2021"
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
library(modelr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```


## Problem 1

**Load and clean data**

```{r,message = FALSE}
birth_weight = 
  read_csv("data/birthweight.csv")
```

First, convert appropriate variables to factors

```{r, message = FALSE}
birth_weight =
  birth_weight %>% 
  mutate(
    babysex = recode_factor(babysex, "1" = "male", "2" = "female"),
    frace = recode_factor(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unkown"),
    mrace = recode_factor(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other"),
    malform = recode_factor(malform, "0" = "absent", "1" = "present" )
  )
```

Summary of the data set before further analysis

```{r}
skimr::skim(birth_weight) %>% 
  select(-c(factor.ordered, factor.n_unique, factor.top_counts,numeric.hist)) %>% 
  knitr::kable()
```

As we can observe from the table, there's no missing value in the data set. Moreover, `pnumlbw` and `pnumgsa` are 0 for all of their observations, therefore those variables should be not used in the regression analysis. 

**Hypothesized Regression Model**

Hypothesize a model using the step-wise approach that we learned in Biostatistical Method I! 

Here, I used the step-wise regression function in R. The process is to start with a model using all predictors and then use the backward elimination process to search through models from full to null to find the model with the smallest AIC value.


```{r}
mod_1 = lm(bwt ~ ., data = birth_weight) %>% 
  step(direction = "backward", trace = 0)

mod_1
```

So, the first model that I hypothesize with backward elimination include 11 variables. Those variables are `babysex`,`bhead`,`blength`,`delwt`,`fincome`, `gaweeks`,`mheight`, `mrace`, `parity`, `ppwt`, and `smoken`.

```{r}
summary(mod_1) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>%
  knitr::kable(digits = 3)
```

Plot of model residuals against fitted values

```{r}
birth_weight %>% 
  add_predictions(mod_1) %>% 
  add_residuals(mod_1) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(se = FALSE) +
  labs(x = "Fitted Values", y = "Residuals",title = "Residuals vs Fitted 1")
```

From the plot, we can see that most of the residuals bounce around 0 in the (-1000,1000) horizontal band. However, there are some concerning points at the lower end of the plot with high residuals. 

**Compare models**

```{r}
mod_2 = lm(bwt ~ blength + gaweeks, data = birth_weight)
mod_3 = lm(bwt ~ bhead + blength + babysex + 
             bhead * blength + 
             bhead * babysex + 
             blength * babysex + 
             bhead * blength * babysex, data = birth_weight)
```

Residual vs Fitted plot for the second model

```{r}
birth_weight %>% 
  add_predictions(mod_2) %>% 
  add_residuals(mod_2) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(se = FALSE) +
  labs(x = "Fitted Values", y = "Residuals",title = "Residuals vs Fitted 2")
```

Residual vs Fitted plot for the second model

```{r}
birth_weight %>% 
  add_predictions(mod_3) %>% 
  add_residuals(mod_3) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(se = FALSE) +
  labs(x = "Fitted Values", y = "Residuals",title = "Residuals vs Fitted 3")
```

Cross Validation


